---
title: "Problem Set 2 Scratch"
output: html_notebook
---

For this problem set, I will be fixing and analyazing a set of Halloween themed data.  The first thing that we will do is clear R's environment and import the correct libraries
```{r}
# clear the environment of R
rm(list = ls())

# import the libraries that we will need
library(ggplot2)
library(dplyr)
library(here)
```

Great, off to a good start so far.  Let's move on to import the data that we will need.  When I open this, I will use the `stringsAsFactors = TRUE` function to convert categorical data into factor data.  This allows R to interpret this as factors with different levels that correspond to the different categories within the factor. 

```{r}
# import the csv file that we are going to use for this exercise without stringsAsFactors = TRUE
Halloween <- read.csv(here("Data", "Halloween_data_PS2.csv"), stringsAsFactors = TRUE)

# open the dataset
View(Halloween)

# now we are going to run a summary of glimpse to take a look at the data
glimpse(Halloween)
```
Ok, still cranking so far - let's keep going...

The next task is to examine the dataset with at least 2 functions.  I used the `glimpse()` function either to get a brief summary of the data.  Now we are going to use the `str()` functions and the `summary()` functions to take a look into some basic statistics regarding the dataset.

```{r}
# get basic stats from the dataset
summary(Halloween)

# get more info surrounding the structure of the dataset
str(Halloween)
```
Now on to fixing errors in the data set.  As of right now, there are a few errors with how the data was inputted in the categorical columns.  For example, right now R believes the costume column has 15 levels but if we look at it using the `View()` function, we can see that there really are 10 categories.  The reason R believes there are 15 is because there are 15 different ways that the data are formatted.  R is case sensitive so the way that the data were formatted does matter. 

To fix the "costume" category, I will reassign the data that refer to the same category to the same name to make the formatting consistent. 
```{r}
# reassign the data that refers to the wonder woman costume as "Wonder Woman"
WW <- which(Halloween$Costume == "Wonder Woman" | Halloween$Costume == "Wonder woman" | Halloween$Costume == "wonder woman")

Halloween$Costume[WW] <- "Wonder Woman"

# reassign the data that refers to the clown costumes as "clown". I am choosing to include the "creepy clown" entries in this as well.  While this does generalize the data a little more, I believe that simplifying the data will make it more digestible. 
clown <- which(Halloween$Costume == "Creepy Clown" | Halloween$Costume == "clown" | Halloween$Costume == "creepy clown" | Halloween$Costume == "Creepy clown")

Halloween$Costume[clown] <- "clown"

# reassign the data that refers to the witch costumes as "witch"
witch <- which(Halloween$Costume == "witch" | Halloween$Costume == "Witch" )

Halloween$Costume[witch] <- "witch"
```

