---
title: "Problem Set 2 Scratch"
output: html_notebook
---

For this problem set, I will be fixing and analyazing a set of Halloween themed data.  The first thing that we will do is clear R's environment and import the correct libraries
```{r}
# clear the environment of R
rm(list = ls())

# import the libraries that we will need
library(ggplot2)
library(dplyr)
library(here)
```

Great, off to a good start so far.  Let's move on to import the data that we will need.  When I open this, I will use the `stringsAsFactors = TRUE` function to convert categorical data into factor data.  This allows R to interpret this as factors with different levels that correspond to the different categories within the factor. 

```{r}
# import the csv file that we are going to use for this exercise without stringsAsFactors = TRUE
Halloween <- read.csv(here("Data", "Halloween_data_PS2.csv"), stringsAsFactors = TRUE)

# now we are going to run a summary of glimpse to take a look at the data
glimpse(Halloween)
```

